Shell Model for Reconstruction and Real-Time Simulation of Thin Anatomical Structures

### 2010-06-28 MICCAI2010/paper.tex
- Latest modifications for camera-ready MICCAI 2010 - fixed cataract scene (threshold angle to return a null quaternion was changed in the qDiffZ method) - scenes updated
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@121 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\title{Shell Model for Reconstruction and Real-Time Simulation of Thin Anatomical Structures}
\author{Olivier Comas\inst{1,2}, Christian Duriez\inst{1} and St\'ephane Cotin\inst{1}}
\institute{INRIA, Shaman team, Lille, France
\and The Australian e-Health Research Centre, CSIRO, Brisbane, Australia}

@@

The human body is composed of various deformable anatomical structures. 
%Realistic modelling of organ deformations is a challenging research field that opens the door to new clinical applications including: medical training and rehearsal systems, patient-specific planning of surgical procedure and per-operative guidance based on simulation. In all these cases the clinician needs fast update of the deformation model to obtain a real-time display of the computed deformations. 
%Another 
A key challenge of soft-tissue modelling is the variousness of the mechanical behaviours. It seems unrealistic to use a unique model for all tissues.
Yet most of previous works focus on volumetric models that are able to capture the behaviour of solid organs like the liver or the brain (see for instance \cite{Miller07,Delingette08}). In contrast, this paper seeks to propose a solution for simulating, in real-time, the deformation of thin anatomical structures whose volume is negligible compared to their surface area. Examples include hollow structures, such as the wall of blood vessels, or membranes, such as the Glisson's capsule surrounding the liver. 
%of such anatomical structures

@@

To model the deformation of complex anatomical structures using shell elements, the first step is to describe its surface with curved patches. This process is quite similar to the reconstruction of the surface of objects in computer vision. Indeed calculating curvature maps of 3D surfaces represented as digitised data (point clouds or triangulated meshes) has been extensively studied. One of the most common approach is to use continuous surface patches~\cite{Kolb95}. The surface is locally approximated by an analytic representation called surface patch, usually chosen to be an implicit surface. These works target approaches that are not noise-sensitive and coherent surface and curve extraction from 3D data~\cite{Tang99}. 
% removed citation Douros02 which was with Kolb95.
%In~\cite{Tang99} an elegant and non noise-sensitive approach that infers sign and direction of principal curvatures directly from the input is presented and the authors use this information for coherent surface and curve extraction from 3D data. 
However, our situation is substantially different as we want to model the deformation of the structure. In that regard the curvature of the surface has a physical meaning: it represents the mid-surface of the shell on which the plane stress hypothesis~\cite{Liu03} applies. We propose to approximate the surface of anatomical structures with shell elements whose each surface is described by the shape function used in our shell formulation.

@@

\subsubsection{Coupling between tetrahedra and shells for advanced modelling.}
%
%Structures in human body can be either solid (brain, liver, prostate etc.) or hollow (blood vessels, colon, stomach etc.). 
Structures in human body can be either solid or hollow. However knowing how to model the two kind of structures is not sufficient to reach a high degree of accuracy, real life situations are more complex. As an example, the external surface of the liver is covered by a layer of cells called Glisson's capsule. Its interaction with the liver plays an important role into the overall structure's mechanical behaviour. Therefore considering the interaction between solid and hollow objects is as crucial as modelling the two structures separately.

@@

\section*{Acknowledgments}
We would like to thank INRIA and the SOFA project, as well as CSIRO Preventative Health National Research Flagship, ICTC, The Australian e-Health Research Centre for their fundings.

>>>>

\title{Shell Model for Reconstruction and Real-Time Simulation of Thin Anatomical Structures}
\author{Olivier Comas\inst{1,2}, Christian Duriez\inst{1} and St\'ephane Cotin\inst{1}}
\institute{INRIA, Shaman team, Lille, France
%\and Preventative Health National Research Flagship, CSIRO ICT, The Australian e-Health Research Centre, Brisbane, Australia}
\and Preventative Health Flagship, CSIRO ICT, AEHRC, Brisbane, Australia}

@@

The human body is composed of various deformable anatomical structures. 
%Realistic modelling of organ deformations is a challenging research field that opens the door to new clinical applications including: medical training and rehearsal systems, patient-specific planning of surgical procedure and per-operative guidance based on simulation. In all these cases the clinician needs fast update of the deformation model to obtain a real-time display of the computed deformations. 
%Another 
A key challenge of soft-tissue modelling is the variousness of the mechanical behaviours. It seems unrealistic to use a unique model for all tissues.
Yet most of previous works focus on volumetric models that are able to capture the behaviour of solid organs like the liver or the brain (see for instance \cite{Miller07,Delingette08}). In contrast, this paper seeks to propose a solution for simulating, in real-time, the deformation of thin anatomical structures whose volume is negligible compared to their surface area. Examples include hollow structures, such as the wall of blood vessels, or membranes, such as the Glisson's capsule surrounding the liver. It is also of particular interest to us for modelling the colon in our colonoscopy simulator \cite{deVisser10}.
%of such anatomical structures

@@

To model the deformation of complex anatomical structures using shell elements, the first step is to describe its surface with curved patches. This process is quite similar to the reconstruction of the surface of objects in computer vision. Indeed calculating curvature maps of 3D surfaces represented as digitised data (point clouds or triangulated meshes) has been extensively studied. One of the most common approach is to use continuous surface patches~\cite{Kolb95}. The surface is locally approximated by an analytic representation called surface patch, usually chosen to be an implicit surface. These works target approaches that are not noise-sensitive and coherent surface and curve extraction from 3D data~\cite{Tang99}. 
% removed citation Douros02 which was with Kolb95.
%In~\cite{Tang99} an elegant and non noise-sensitive approach that infers sign and direction of principal curvatures directly from the input is presented and the authors use this information for coherent surface and curve extraction from 3D data. 
However, our situation is substantially different as we want to model the deformation of the structure. In that regard the curvature of the surface has a physical meaning: it represents the mid-surface of the shell. 
% on which the plane stress hypothesis~\cite{Liu03} applies. 
We propose to approximate the surface of anatomical structures with shell elements whose each surface is described by the shape function used in our shell formulation.

@@

\subsubsection{Coupling between tetrahedra and shells for advanced modelling.}
%
%Structures in human body can be either solid (brain, liver, prostate etc.) or hollow (blood vessels, colon, stomach etc.). 
Structures in human body can be either solid (brain, liver, prostate etc.) or hollow (colon, blood vessels, stomach etc.). However knowing how to model the two kind of structures is not sufficient to reach a high degree of accuracy, real life situations are more complex. As an example, the external surface of the liver is covered by a layer of cells called Glisson's capsule. Its interaction with the liver plays an important role into the overall structure's mechanical behaviour. Therefore considering the interaction between solid and hollow objects is as crucial as modelling the two structures separately.

@@

%\section*{Acknowledgments}
%We would like to thank INRIA and the SOFA project, as well as CSIRO Preventative Health National Research Flagship, ICTC, The Australian e-Health Research Centre for their fundings.
